---
title: "P8105_hw2_jf3286.Rmd"
author: "Jingyu Fu"
date: "2019/9/26"
output: github_document
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE,
                      fig.width = 6,
                      fig.asp = .6,
                      out.width = "90%")
library(tidyverse)
library(dplyr)

```

# Problem 1

## Read and clean the Mr. Trash Wheel sheet 
```{r}
library(readxl)
Mr_Trash_Wheel_data = read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", sheet = "Mr. Trash Wheel", range = "A2:N408", col_names = TRUE) %>%
  janitor::clean_names()%>%
   rename (total_weight_tons = weight_tons,  total_volume_cubic_yards = volume_cubic_yards)%>%
   drop_na (dumpster) %>% 
   mutate(
    sports_balls = round(sports_balls, digits = 0),
    sports_balls = as.integer(sports_balls)
  ) 
Mr_Trash_Wheel_data
```
    
## Read and clean precipitation data for 2017 and 2018. 
```{r}
 precipitation_2017 = read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", sheet = "2017 Precipitation", skip = 1, col_names = TRUE) %>% 
    janitor::clean_names() %>% 
    drop_na ()%>%
   mutate (year = 2017) %>% 
   select (year,everything())
 
 precipitation_2018 = read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", sheet = "2018 Precipitation", skip = 1, col_names = TRUE) %>% 
    janitor::clean_names()%>%
    drop_na ()%>%
   mutate (year = 2018) %>% 
   select (year,everything())
 precipitation_2017
```
  
## Next, combine precipitation datasets. 
```{r} 
 combination_precipitation = full_join (precipitation_2017,precipitation_2018) %>% 
   mutate (
     month = as.numeric (month),
     month = month.name [month]
     
   )
combination_precipitation
``` 

## Write a paragraph about these data.

The total precipitation in 2018 is `r sum(pull(precipitation_2018, total))`.The median number of sports balls in a dumpster in 2017 is `r median(pull(filter(Mr_Trash_Wheel_data, year == 2017), sports_balls))`. In Mr_Trash_Wheel_data, there are `r nrow(Mr_Trash_Wheel_data)`rows and `r ncol(Mr_Trash_Wheel_data)`columns, and the key variable is dumpster, which identifies each row.In combination_precipitation_data,there are `r nrow(Mr_Trash_Wheel_data)`rows and `r ncol(Mr_Trash_Wheel_data)`columns, and the key variable is month, which is the link between precipitation 2017 data and precipitation 2018 data. In precipitation_2017 data, there are `r nrow(precipitation_2017)`rows and `r ncol(precipitation_2017)`columns.The mean  is `r mean(pull(precipitation_2017, total))`.The key variable is month, because the month identifies each row.In the precipitation_2018 data, there are `r nrow(precipitation_2018)`rows and `r ncol(precipitation_2018)`columns. The mean is `r mean(pull(precipitation_2018, total))`.The key variable is month, because the month identifies each row.

 


# Problem 2

## First, clean the data in pols-month.csv.
```{r}
pols_month = read_csv(file = "./data/fivethirtyeight_datasets/pols-month.csv", col_names = TRUE)%>%
  janitor:: clean_names () %>% 
  separate (mon, into = c("year", "month", "day"), convert=T) %>% 
  mutate(
    month = as.numeric (month), 
    month = month.name [month],
    president = ifelse (prez_gop == 0, "dem","gop")
  ) %>% 
  select(- prez_dem, - prez_gop, -day)
pols_month
```
 
## Second, clean the data in snp.csv. 
```{r}
snp = read_csv(file = "./data/fivethirtyeight_datasets/snp.csv") %>% 
janitor:: clean_names () %>% 
  separate (date, into = c( "month", "day","year"), convert=T) %>% 
  arrange (year, month) %>% 
  select (year, month, everything (),-day) %>% 
  mutate(
    month = as.numeric (month), 
    month = month.name [month]
  ) 
snp
  
```


## Third, tidy the unemployment data.  
```{r}
unemployment = read_csv(file = "./data/fivethirtyeight_datasets/unemployment.csv") %>% 
  pivot_longer (
  Jan:Dec, 
  names_to = "month",
  values_to = "number"
              ) %>% 
janitor:: clean_names () 
unemployment
```


## Join the datasets. 
```{r}
join_datasets = 
  left_join (pols_month, snp, by = c ("year","month")) %>% 
    left_join (unemployment,by = c ("year","month"))
join_datasets
```

## Write a short paragraph about these datasets. 

In pols-month.csv , there contained the date, and corresponding information about president government. For its resulting dataset, there are `r nrow(pols_month)` rows and `r ncol(pols_month)` columns, range of year is(`r range(pull(pols_month,year))`), and the key variable is date, which identifies each row. In snp.csv , there contained the date, and corresponding information about president government. For its resulting dataset, there are `r nrow(snp)` rows and `r ncol(snp)` columns, range of year is(`r range(pull(snp,year))`), and the key variables are year, month and day, which together identifies each row. In unemployment.csv , there contained the date, and corresponding information about president government. For its resulting dataset, there are `r nrow(unemployment)` rows and `r ncol(unemployment)` columns, range of year is(`r range(pull(unemployment,year))`), and the key variables are year and month, which together identify each row. 


# Problem 3

## Load and tidy the data.
```{r}
library(ggridges)
library(dplyr)
popular_baby_names = read_csv (file = "./data/Popular_Baby_Names.csv", col_names = TRUE, col_types = "ncccnn")%>%
  janitor::clean_names()%>%
   mutate(
    ethnicity = str_to_lower(ethnicity),
    childs_first_name = str_to_lower(childs_first_name),
    gender = str_to_lower(gender),
    ethnicity = recode(ethnicity,
                       "asian and paci" = "asian and pacific islander",
                       "black non hisp" = "black non hispanic",
                       "white non hisp" = "white non hispanic")
    
  )  %>% 
distinct(.keep_all = TRUE)
popular_baby_names
```
 
## Produce a table showing the rank in popularity of the name “olivia” as a female baby name over time.
```{r}
  popular_baby_names_1 = 
    select (popular_baby_names, - count)%>%
  filter (gender == "female" & childs_first_name == "olivia")%>%
  pivot_wider(names_from = "year_of_birth", values_from = "rank")%>%
  select (- gender, - childs_first_name)
  knitr::kable(popular_baby_names_1)
```

## Produce a similar table showing the most popular name among male children over time.
```{r}
 popular_baby_names_2 =
  filter (popular_baby_names, gender %in% c ("male") & rank == 1 )%>%
  select (- count, - gender)%>%
  pivot_wider(names_from = "year_of_birth", values_from = "childs_first_name") 
 
  knitr::kable(popular_baby_names_2)
```
                                                                                    
## Finally, produce a scatter plot for male, white non-hispanic children born in 2016.                                                                               
```{r}
 ggplot_data = 
   filter (popular_baby_names, gender =="male"& ethnicity == "white non hispanic" & year_of_birth == 2016) 

  ggplot(ggplot_data, aes(x = rank, y = count)) + 
    geom_point()+
 labs(
    title = "Name and rank plot  ",
    x = "Rank in popularity of a name",
    y = "Number of children with a name ",
    caption = "Data for white non-hispanic male children born in 2016"
  )
  

```
